Light inextensible strings under tension in the Schwarzschild geometry

### 2019-08-29 string_physics_teacher.tex
clarification

There does not appear to be a simple analytical solution to
equation~\ref{rdashdash} but it is a straightforward second order ODE,
amenable to numerical solution.  Here we consider only strings
exterior to the event horizon, $r>2M$, and for these we have a
well-defined point of closest approach to the event horizon.  Further,
because the system is invariant under~$\phi\longrightarrow-\phi$, the
string must be symmetrical about at this point.
Considering~$r'(\phi)=0$ shows that a tangential string at a point
with~$r_0>2M$ obeys, to first order, $r''=r-2M$, while tangential null
geodesics follow $r''=r-3M$.  Thus a taut string would {\em look}
curved to a local observer, having a (visual) radius of curvature of
$r_0^2/M\sqrt{1-2M/r_0}$ (the extra term comes from the fact that
appearance is driven by local $ds$, not Schwarzschild $r$).  At the
photon sphere $r_0=3M$, such a string would appear to have a radius
of~$\sqrt{27}M$.  For a solar mass black hole this would be about
$15\unit{km}$, a readily understandable result.

@@

\section*{References}
\bibliography{stringrefs.bib}

>>>>

There does not appear to be a simple analytical solution to
equation~\ref{rdashdash}; but it is a straightforward second order
ODE, amenable to numerical solution and no harder than the equivalent
for null geodesics, a standard homework problem.  Here we consider
only strings exterior to the event horizon, $r>2M$, which have a
well-defined point of closest approach to the singularity.  Because
the system is invariant under~$\phi\longrightarrow-\phi$, the string
must be symmetrical about this point.  If $\phi'$ is small, then a
string follows $r''=r-2M$, while tangential null geodesics follow
$r''=r-3M$.  Thus a taut string would {\em look} curved to a local
observer, having a (visual) radius of curvature of
$r_0^2/M\sqrt{1-2M/r_0}$ (the extra term comes from the fact that
appearance is driven by local $ds$, not Schwarzschild $r$).  At the
photon sphere $r_0=3M$, such a string would appear to have a radius
of~$\sqrt{27}M$.  For a solar mass black hole this would be about
$15\unit{km}$, a readily understandable result.

@@

\bibliography{stringrefs.bib}
